---
title: "Individual Assignment 5"
---


Load packages and data; we will be using the vegan library to calculate similarity measures
```{r}

library(pacman)
p_load(tidyverse,arules)

#read in the patient data
Members <- read_csv("members.csv")
#read in the diagnosis descriptions
DiagDesc <- read_csv("DiagDictionary.csv")

```

Task 1: summarizing diagnosis (Feel free to complete this task in Excel if you prefer)
```{r}
#Let's work with just the diagnosis data
DiagData<-Members[,5:51]

#get the corresponding description
DiagDesc<-DiagDesc[5:51,3]

#now summarize the data
DiagCounts = apply(DiagData,2,sum)

```



TASK 2 - Optional Data Reduction Step: Reduce the data while testing the code and getting to know how much data your machine can handle. 
TASK 2 - Using threshold to filter member count
```{r}
#Let's work with just the diagnosis data
DiagData<-Members[,5:51]

#set this number to some low number depending on how powerful your computer is
MinDiag=3

#count the diagnosis for each member
DiagData <- DiagData %>% rowwise %>% mutate(DiagCount=sum(across(where(is.numeric))))

#filter the dataset
DiagData <- DiagData %>% filter(DiagCount>=MinDiag)

#drop the DiagCount
DiagData <-DiagData %>% select(-DiagCount)
```



Task 2: Run the algorithm on the data created

Lets start with transforming the data
```{r}
#Removing Diag_DD0046
DiagData<-DiagData %>% select(-Diag_DD0046)
DiagDataDF <- DiagData
# convert the binary incidence matrix into a transactions database
DiagData<-as.matrix(DiagData)
fp.trans <- as(DiagData, "transactions")

#use inspect to verify that the data looks as expected
#inspect(fp.trans)

```
Task 3: run rules on customized supp and conf levels
```{r}
rules <- apriori(fp.trans, parameter = list(supp = 0.01, conf = 0.1, target = "rules"))
inspect(head(sort(rules, by = "lift"), n = 3))
```

OPTIONAL: loop to find highest lift
```{r}
for( i in seq(0.01,1,0.05)){
  for (ii in seq(0.01,1,0.05)) {
    loop.rules <- apriori(fp.trans, parameter = list(supp = i, conf = ii, target = "rules"))
    loop.list<- c(i,ii,rules@quality[4])
  }
}
loop.list
```

Save your rules to a file

```{r}
#rules is the output of the association rule algorithm
write(rules, file = "association_rules.csv", sep = ",", quote = TRUE, row.names = FALSE)
rule_data <- DATAFRAME(rules, separate = TRUE, setStart = '', itemSep = ',', setEnd = '')
write.csv(rule_data,"MyRules.csv")
```

question 8
```{r}
Members<-Members %>%
  mutate(TA2 = Members['MedPaidAmount'] + Members['RxPaidAmount'], HighCost = ifelse(TA2>=30000,1,0))

DiagData.HC<-cbind(Members[,5:51],Members[,242])

MinDiag=3

DiagData.HC <- DiagData.HC %>% rowwise %>% mutate(DiagCount=sum(across(where(is.numeric))))

DiagData.HC<- DiagData.HC %>% filter(DiagCount>=MinDiag)

DiagData.HC <-DiagData.HC %>% select(-DiagCount)

DiagData.HC<-DiagData.HC %>% select(-Diag_DD0046)

DiagData.HC<-as.matrix(DiagData.HC)
fp.trans.HC <- as(DiagData.HC, "transactions")

```
question 8 - run rules & output
```{r}
rules.HC <- apriori(fp.trans.HC, parameter = list(supp = 0.001, conf = 0.1, target = "rules"),appearance = list(rhs = "HighCost"))
inspect(head(sort(rules.HC, by = "lift"), n = 5))

write(rules.HC, file = "association_rules_HC.csv", sep = ",", quote = TRUE, row.names = FALSE)
rule_data.HC <- DATAFRAME(rules.HC, separate = TRUE, setStart = '', itemSep = ',', setEnd = '')
write.csv(rule_data.HC,"MyRules_HC.csv")
```


